---
title: "map_dbmart_ccsr_act"
author: "Jiazi Tian"
date: "6/26/2023"
output: html_document
---

```{r load packages}
library(dplyr)
library(stringr)
library(data.table)
```

# User input
The user needs to manually enter the group, group data directory, the output directory and site name.
```{r user input}
###### Modify the following lines
##utils::choose.dir is a windows functionality use tk on other systems
choose_directory = function(caption = 'Select directory') {
  if(commandArgs()[[1L]] == "RStudio"){
    rstudioapi::selectDirectory(caption = caption)
  }else if (exists('utils::choose.dir')) {
    choose.dir(caption = caption)
  } else{
    tcltk::tk_choose.dir(caption = caption)
  }
}

###### Modify the following lines
group <- "cases" #cases, controls, controls_pre

groupDirectory <- choose_directory(caption = "select group data directory")  ## where the inputs come from 
outputDirectory <- choose_directory(caption = "select output data directory") ## where the outputs are saved
site <- "MGH" 
######
```


# Load data
```{r load data}
input_file <- paste0(groupDirectory, "/",group, ".csv")
df <-  data.table::fread(input_file) 
ccsr <- data.table::fread("CCSR_PASC_ACT_Mapping_part_082023.csv")
```


# Preprocess the data

## Extract columns from the input file
This step is to extract the target four columns, 'PATIENT_NUM', 'START_DATE', 'CONCEPT_CD' and 'C_FULLNAME', from the input file for the mapping process. If your input file doesn't contain these four columns then it won't work and you need to check your data.
```{r extract columns}
names(df) <- toupper(names(df))
# Patient data should have at least these four columns
target_cols <- c("PATIENT_NUM", "START_DATE", "CONCEPT_CD","C_FULLNAME")

if (all(target_cols %in% colnames(df))) {
  df <- subset(df, select = target_cols)
} else {
  print("The input data doesn't include 'PATIENT_NUM', 'START_DATE', 'CONCEPT_CD' or 'C_FULLNAME' columns.")
}
```


## Remove the duplicates
This step is to remove the duplicates, and create a domain column based on the paths.
```{r remove duplicates}
# remove duplicates, create domain column
df_rmdup <- unique(df)
df_rmdup$domain <- str_extract(df_rmdup$C_FULLNAME, "^\\\\[^\\\\]+\\\\[^\\\\]+\\\\[^\\\\]+")
rm(df)
```

## Cases check
If the input data is cases,  COVID positive indications are extracted from the data and saved as cov_pats.RData. Otherwise, continue to the next step. 
```{r}
if(group == 'cases'){
  rdx <-  df_rmdup[(df_rmdup$CONCEPT_CD=="ICD10:U07.1"), ]
  rp1 <-  df_rmdup[startsWith(df_rmdup$C_FULLNAME, "\\ACT\\UMLS_C0031437\\SNOMED_3947185011\\UMLS_C0037088\\SNOMED_3947183016\\"), ]
  rp2 <-  df_rmdup[startsWith(df_rmdup$C_FULLNAME, "\\ACT\\UMLS_C0031437\\SNOMED_3947185011\\UMLS_C0022885\\UMLS_C1335447\\"), ]
  rp3 <- df_rmdup[startsWith(df_rmdup$C_FULLNAME, "\\ACT\\UMLS_C0031437\\SNOMED_3947185011\\UMLS_C0022885\\ACT_LOCAL_LAB_ANY_POSITIVE\\"), ] 
  cases_encs <- rbind(rdx, rp1, rp2, rp3)
  save(cases_encs, file= paste0(groupDirectory, "/cov_pats.RData")) 
  
  df_rmdup <- anti_join(df_rmdup, cases_encs)
 
  rm(rdx, rp1, rp2, rp3, cases_encs)
}else{
  print("Continue to the next step.")
}
```



## Extract Diagnosis_ICD10, Procedures_ICD10 and UMLS ICD10
This step is to obtain data for the Diagnosis_ICD10, Procedures_ICD10 groups and UMLS ICD10, and take a quick look at the number of unique codes and the total number of records.
```{r extract ICD10}
df_ICD10 <- subset(df_rmdup, domain %in% c("\\ACT\\Diagnosis\\ICD10", "\\ACT\\Procedures\\ICD10"))
df_umls_cd <- subset(df_rmdup, domain == "\\ACT\\UMLS_C0031437\\SNOMED_3947185011" & startsWith(CONCEPT_CD, "ICD10"))
df_cd <- rbind(df_umls_cd, df_ICD10)
rm(df_ICD10, df_umls_cd)
smr_df_ICD10 <- df_cd %>% group_by(domain) %>% summarise(nunique=n_distinct(CONCEPT_CD), total = n())
smr_df_ICD10
```

# Merge based on the path

## Step 1: Direct merge based on the paths
We directly use the path as a key to map CCSR to input data and obtain the mapping success rate. If the mapping success rate is less than 0.9, go to step 2 to modify the path. Otherwise, go to the final step to summarize the results and then output the results.
```{r}
df_cd$CONCEPT_PATH <- df_cd$C_FULLNAME
map_path <- merge(df_cd, ccsr, by = "CONCEPT_PATH", all.x = TRUE)

ms_rate <- n_distinct(map_path$CONCEPT_CD[!is.na(map_path$CCSR_Key)])/n_distinct(map_path$CONCEPT_CD)
ms_rate
```



### If the mapping success rate is less than 0.9, go to step 2. Otherwise, go to summarize the results.
```{r}
if (ms_rate < 0.9){
  df_1_mapped <- map_path[!is.na(map_path$CCSR_Key),]
  df_2 <- subset(map_path, is.na(CCSR_Key), select= c("PATIENT_NUM", "START_DATE", "CONCEPT_CD","C_FULLNAME", "domain"))
  print("Go to step 2.")
  rm(map_path)
} else {
  df_1_mapped <- map_path[!is.na(map_path$CCSR_Key),]
  print("Go to summarize the results.")
  rm(map_path)
}
```


## Step 2: Modify the path by removing the last child
We modify the path by removing the last child from the path. However, there may be different ways to modify the path. Users can change the following pattern by themselves.
```{r}
pattern <- "\\\\[^\\\\]+\\\\?$"
df_2$CONCEPT_PATH <- gsub(pattern, "\\\\", df_2$C_FULLNAME)
```

### Merge based on the paths
After modifying the paths, we use the new paths again to map CCSR to input data and obtain the new mapping success rate. Then we check the mapping success rate again, if it is still less than 0.9,the user needs to manually compare the path with the ACT standard path.
```{r}
map_path_2 <- merge(df_2, ccsr, by = "CONCEPT_PATH", all.x = TRUE)
ms_rate_2 <- n_distinct(map_path_2$CONCEPT_CD[!is.na(map_path_2$CCSR_Key)])/n_distinct(map_path_2$CONCEPT_CD)
ms_rate_2
```


### If the mapping success rate is still less than 0.9, manually check the path in the input file or change the pattern in step 2. Otherwise, go to summarize the result.
```{r}
if ((ms_rate_2) < 0.9){
  print("Need to manually check the path")
} else {
  df_2_mapped <- map_path_2[!is.na(map_path_2$CCSR_Key),]
  print("Go to summarize the results.")
  rm(map_path_2)
}
```


# Summarize the results

## Statistics on successful mappings
Here we report statistics on successful mappings, including the number of successfully mapped unique codes, the number of successfully mapped records, the percentage of successfully mapped codes, and the percentage of successfully mapped records.
```{r}
if (exists("df_2_mapped")){
  mapped <- rbind(df_1_mapped, df_2_mapped)
  rm(df_2_mapped)
} else{
  mapped <- df_1_mapped
}
rm(df_1_mapped)

smr_mapped <- mapped %>% group_by(domain) %>% summarise(nunique=n_distinct(CONCEPT_CD), total=n())

result <-  cbind(smr_mapped, round(smr_mapped[-1]/smr_df_ICD10[-1]*100, digits = 3))
names(result)[-1] <- c("mapped_unique_codes", "mapped_total_records", "mapped_unique_codes_percentage", "mapped_total_records_percentage")
result
```

## Demographics and race summary
```{r}
dems_all <-  data.table::fread(paste0(groupDirectory, "/dems_",group, ".csv"))
dems <- dems_all[dems_all$patient_num %in% mapped$PATIENT_NUM, c("patient_num", "age", "sex_cd", "CHARLSON_INDEX", "race_cd")]
# When the patient has more than 1 ages, select the oldest.
dems <- unique(dems) %>% group_by(patient_num) %>% 
  filter(age == max(age)) %>% 
  ungroup()
# When the patient has more than CHARLSON INDEX, select the largest.
dems <- unique(dems) %>% group_by(patient_num) %>% 
  filter(CHARLSON_INDEX == max(CHARLSON_INDEX)) %>% 
  ungroup()

dat.sum <- subset(mapped,mapped$PATIENT_NUM %in% dems$patient_num)
colnames(dat.sum) <- tolower(colnames(dat.sum))

dems_temp1 <- data.frame(rbind(c("patients",nrow(dems))))
dems_temp2 <- data.frame(rbind(c("mean age",mean(dems$age))))
dems_temp3 <- data.frame(rbind(c("percent female",(nrow(subset(dems,dems$sex_cd == "F"))/nrow(dems))*100)))

dems_temp4 <- data.frame(rbind(c("mean charlson",mean(dems$CHARLSON_INDEX))))

tb1 <- rbind(dems_temp1,dems_temp2,dems_temp3,dems_temp4) 
tb2 <- data.frame(rbind(c("unique CCSR concepts",length(unique(dat.sum$phenx)))))
length.all <- dat.sum %>% group_by(patient_num) %>% summarise(max=max(start_date),min=min(start_date))

length.all$durate=as.numeric(difftime(length.all$max, length.all$min,unit="weeks"))/52.25

tb3 <- data.frame(rbind(c("mean data depth",mean(length.all$durate))))
tb4 <- data.frame(rbind(c("data to",max(length.all$max))))
tb5 <- data.frame(rbind(c("data from",min(length.all$min))))

table1 <- rbind(tb1,tb2,tb3,tb4,tb5)
table1$site <- site
table1$group <- group

race <- data.frame(table(dems$race_cd))
race <- subset(race,race$Freq > 30)
race$site <- site
race$group <- group
```


# Output the results
Four files, the demographics, race and the statistics on successful mappings are saved in the outputs directory, and the mapped result file are saved in the group data directory.
```{r}
fwrite(table1, paste0(outputDirectory,"/", group,"_dems_", site, ".csv" ))
fwrite(race, paste0(outputDirectory, "/",group,"_race_", site, ".csv" ))
fwrite(result, paste0(outputDirectory, "/",group,"_map_stats_", site, ".csv" ))
fwrite(mapped, paste0(groupDirectory,"/", group, "_map_CCSR_",site, ".csv"))
```


