---
title: "map_dbmart_ccsr_act"
author: "Jiazi Tian"
date: "6/26/2023"
output: html_document
---

```{r load packages}
library(dplyr)
library(stringr)
```

# User input
The user needs to manually enter the input file path, data frame name, output directory and site name.
```{r}
###### Modify the following lines
group <- "cases"
output_dict <- paste0("~/clai_share/shared_workspace/ACT/data/",group) ##where the output file is saved
df <- data.table::fread(file.choose()) ###input data
site <- "MGH" 
######
```


# Load data
```{r load data}
ccsr <- data.table::fread(file.choose()) ###CCSR mapping file
```


# Preprocess the data

## Clean CCSR mapping
This step is to remove the data without ccsr key and to extract the target columns, "CONCEPT_PATH", "CCSR_Key","phenx","phenx1","phenx2","map", "relevance" for further analysis.
```{r clean CCSR}
ccsr_m <- subset(ccsr, CCSR_Key!= "", select = c("CONCEPT_PATH", "CCSR_Key","phenx","phenx1","phenx2","map", "relevance"))
rm(ccsr)
```

## Extract columns from the input file
This step is to extract the target four columns, 'PATIENT_NUM', 'START_DATE', 'CONCEPT_CD' and 'C_FULLNAME', from the input file for the mapping process. If your input file doesn't contain these four columns then it won't work and you need to check your data.
```{r extract columns}
names(df) <- toupper(names(df))
# Patient data should have at least these four columns
target_cols <- c("PATIENT_NUM", "START_DATE", "CONCEPT_CD","C_FULLNAME")

if (all(target_cols %in% colnames(df))) {
  df <- subset(df, select = target_cols)
} else {
  print("The input data doesn't include 'PATIENT_NUM', 'START_DATE', 'CONCEPT_CD' or 'C_FULLNAME' columns.")
}
```


## Remove the duplicates
This step is to remove the duplicates, and create a domain column based on the paths. Then, if there are duplicates in the first three columns, we keep the one having the domain starting with "ACT".
```{r remove duplicates}
# remove duplicates, create domain column
df_rmdup <- unique(df) %>%
  mutate(domain = sapply(strsplit(C_FULLNAME, "\\\\"), function(x) paste(x[2:4], collapse = "\\"))) %>%
  group_by(PATIENT_NUM, START_DATE, CONCEPT_CD) %>%
  filter(startsWith(domain, "ACT"))
rm(df);gc()
```

## Extract Diagnosis_ICD10 and Procedures_ICD10
This step is to obtain data for the Diagnosis_ICD10 and Procedures_ICD10 groups and take a quick look at the number of unique codes and the total number of records.
```{r extract ICD10}
df_ICD10 <- subset(df_rmdup, domain %in% c("ACT\\Diagnosis\\ICD10", "ACT\\Procedures\\ICD10"))
smr_df_ICD10 <- df_ICD10 %>% group_by(domain) %>% dplyr::summarise(nunique=n_distinct(C_FULLNAME), total =n())
smr_df_ICD10
```

# Merge based on the path

## Step 1: Direct merge based on the paths
We directly use the path as a key to map CCSR to input data and obtain the mapping success rate. If the mapping success rate is less than 0.9, go to step 2 to modify the path. Otherwise, go to the final step to summarize the results and then output the results.
```{r}
df_ICD10$CONCEPT_PATH <- df_ICD10$C_FULLNAME
map_path <- merge(df_ICD10, ccsr_m, by = "CONCEPT_PATH", all.x = TRUE)

ms_rate <- n_distinct(map_path$C_FULLNAME[!is.na(map_path$CCSR_Key)])/n_distinct(map_path$C_FULLNAME)
ms_rate
```



### If the mapping success rate is less than 0.9, go to step 2. Otherwise, go to summarize the results.
```{r}
if (ms_rate < 0.9){
  df_1_mapped <- map_path[!is.na(map_path$CCSR_Key),]
  df_2 <- subset(map_path, is.na(CCSR_Key), select= c("PATIENT_NUM", "START_DATE", "CONCEPT_CD","C_FULLNAME", "domain"))
  print("Go to step 2.")
  #rm(map_path)
} else {
  df_1_mapped <- map_path[!is.na(map_path$CCSR_Key),]
  print("Go to summarize the results.")
  #rm(map_path)
}
```


## Step 2: Modify the path by removing the last child
We modify the path by removing the last child from the path. However, there may be different ways to modify the path. Users can change the following pattern by themselves.
```{r}
pattern <- "\\\\[^\\\\]+\\\\?$"
df_2$CONCEPT_PATH <- gsub(pattern, "\\\\", df_2$C_FULLNAME)
```

### Merge based on the paths
After modifying the paths, we use the new paths again to map CCSR to input data and obtain the new mapping success rate. Then we check the mapping success rate again, if it is still less than 0.9,the user needs to manually compare the path with the ACT standard path.
```{r}
map_path_2 <- merge(df_2, ccsr_m, by = "CONCEPT_PATH", all.x = TRUE)
ms_rate_2 <- n_distinct(map_path_2$C_FULLNAME[!is.na(map_path_2$CCSR_Key)])/n_distinct(map_path_2$C_FULLNAME)
ms_rate_2
```


### If the mapping success rate is still less than 0.9, manually check the path in the input file or change the pattern in step 2. Otherwise, go to summarize the result.
```{r}
if ((ms_rate_2) < 0.9){
  print("Need to manually check the path")
} else {
  df_2_mapped <- map_path_2[!is.na(map_path_2$CCSR_Key),]
  print("Go to summarize the results.")
}
```


# Summarize the results
Here we report statistics on successful mappings, including the number of successfully mapped unique codes, the number of successfully mapped records, the percentage of successfully mapped codes, and the percentage of successfully mapped records.
```{r}
if (exists("df_2_mapped")){
  mapped <- rbind(df_1_mapped, df_2_mapped)
  rm(df_2_mapped)
} else{
  mapped <- df_1_mapped
}
rm(df_1_mapped)

smr_mapped <- mapped %>% group_by(domain) %>% dplyr::summarise(nunique=n_distinct(C_FULLNAME), total =n())

result <-  cbind(smr_mapped, round(smr_mapped[-1]/smr_df_ICD10[-1]*100, digits = 3))
names(result)[-1] <- c("mapped_unique_codes", "mapped_total_records", "mapped_unique_codes_percentage", "mapped_total_records_percentage")
result
```

# Output the results
Two files, the statistics on successful mappings and the mapped result file, are generated in the output directory.
```{r}
output_stats <- paste0(output_dict, df_name,"_stats_", site, ".csv")
output_file <- paste0(output_dict, df_name, "_CCSR_",site, ".csv")

write.csv(result, output_stats, row.names = FALSE)
write.csv(mapped, output_file, row.names = FALSE)
```


